### 2016-01-19 inputs/acknowledgements.tex
adding the latex folder

\begin{acknowledgements}
The creation of the rice data set was financially supported by the Marie Curie European
Reintegration Grant ``NEUTRADAPT''.
The creation of the alfalfa data sets was funded by the projects ‘Genomic selection in alfalfa (GENALFA)’ funded by the Italian Ministry of Foreign Affairs and International Cooperation in the framework of the Italy-USA scientific cooperation program, the Italian share of the FP7-ArimNet project ‘Resilient, water- and energy-efficient forage and feed crops for Mediterranean agricultural systems (REFORMA)’ funded by the Italian Ministry of Agricultural and Forestry Policies.
\end{acknowledgements}