### 2016-01-19 inputs/table_descriptive_statistics.tex

\begin{table}
\centering
\caption[Descriptive statistics]{
Descriptive statistics of alfalfa (\emph{Medicago sativa}) 
and rice (\emph{Oryza sativa}) genotyping. MAF is average minor allele frequency; 
p(AA), p(AB) and p(BB) are the proportions of AA, AB and BB genotypes.}
\label{tab:descriptive_statistics}
\begin{tabular}{lccccccc}
\hline\noalign{\smallskip}
\noalign{\smallskip}\hline\noalign{\smallskip}
\textbf{Dataset} & \textbf{No. samples} & \textbf{No. SNPs} & \textbf{Missing Rate} & \textbf{MAF} & \textbf{p(AA)} & \textbf{p(AB)} & \textbf{p(BB)}\\
\noalign{\smallskip}\Xhline{3\arrayrulewidth}\noalign{\smallskip}
Alfalfa-PV & 124 & 32\,706 & 0.666 & 0.1724 & 0.690 & 0.274 & 0.035\\
Alfalfa-Med & 154 & 40\,734 & 0.596 & 0.1702 & 0.691 & 0.276 & 0.032\\
Rice & 437 & 166\,418 & 0.534 & 0.140 & 0.860 & - & 0.140 \\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}