### 2016-01-31 reviews/answer_to_revieweres.txt
added answer draft

COMMENTS TO THE AUTHOR:

Reviewer #1: In the MS entitled "Marker imputation efficiency for Genotyping-By-Sequencing data of crop species with and without a reference genome" Authors used different tools to impute missing SNP data in Medicago sativa, and rice. The MS demonstrated very good approach to impute SNP data even independent of whole genome sequence. No doubt the efforts made by authors are valuable. I have some major concern with the MS presentation, flow of information and regarding the novelty. Here are my point-wise comments:

General Comments:
The four methods used for imputation are now very common. Even software tool like Tassel host imputations methods like FILLIN, FSFHap, Numerical, and Beagle. There are several methods and tools available for imputation and many studies demonstrating comparisons have been published.

ANSWER: the main interest in our work was to assess the imputation efficiency in two important plant species, rice (major crop species) and alfalfa (forage species). We wanted to evaluate the effects of increasing missing rates, of aligning markers along the genome, and of the size of the dataset. We were also interested in checking how different imputation methods responded to different conditions. The main points that we think are novel and of interest in our owrk are: i) the imputation of missing data from GBS experiments in alfalfa (M. sativa) and rice (O. sativa); ii) the breakdown of the imputation accuracy into the different genotype classes (major and minor homozygotes, heterozygotes); iii) the effect of aligning markers to the reference genome of the species or to that of a closely related species

More particularly, in species without availability of whole genome sequence methods for the imputation of SNP data have been published. For instance, Ward, et al. (Saturated linkage map construction in Rubus idaeus using genotyping by sequencing and genome-independent imputation. 2013. BMC genomics, 14(1), 2).

Even in Alfaalfa (Medicago sativa) Rocher et al. (Validation of Genotyping-By-Sequencing Analysis in Populations of Tetraploid Alfalfa by 454 Sequencing. 2015. PloS one, 10(6), e0131918.) demonstrated imputation at missing rate 0.5.

ANSWER: the mentioned papers are indeed relevant, and were cited in the introduction of the paper

Specific comments:

In title "crop species with and without" is miss leading since authors have used two species so better to be specific "medicago and rice".

ANSWER: the title was changed to "Marker imputation efficiency for Genotyping-By-Sequencing data in rice (Oryza sativa) and alfalfa (Medicago sativa)"

Abstract need to be more concise and informative

ANSWER:

Introduction:
-------------
Line 38-39, GWAS can not handle missing data- need citation and reframe the sentence. Better to verify the fact.

ANSWER: In the paper, we were actually saying that most models and software for GWAS etc can not handle missing data, which is quite true; GWAS and genomic predictions with missing genotypes is generally more difficult. We added a reference to a review paper by Ben Hayes et al, 2009, in animal genomics. Especially linear algebra and arithmetic methods methods suffer from missing data, whereas numerical methods like MCMC can better cope with them. In any case, using missing data, is not impossible: e.g. maximum likelihood (Handling missing data by maximum likelihood, by P.D. Allison, 2012; or Schafer, 1997: Analysis of incomplete multivariate data). We feel this is out of the scope of the present paper, though, and did not indulge in such speculations. We simply modified the sentence and added a reference.

Line 43-44 - alternative to SNP-chip- need to improve

ANSWER:  The sentence was modified to "Genotyping-by-sequencing (GBS) is a relatively recent technique which is considered a viable alternative to SNP array-based genotyping to produce SNP genotype data"

Line 48 - '52% or up to 52%' need to verify the fact

ANSWER: From Crossa et al: around 52% and 58% in two maize datasets; from Poland et al: quoting from the text "up to 80% missing data per marker" (Page 105)

Line 55 -57 - imputation method has been developed for SNP-Chips. I am not sure about the accuracy of statement

ANSWER: sentence modified to "Most methods for genotype imputation have been applied to SNP array data, and typically yield very high imputation accuracy"

Line 85- rice genome recently assemble - not true

ANSWER: it was assembled in 2005, and then recently updated (2013). We made this clearer in the text

Material methods:
-----------------
Since the MS specifically focused on imputation of SNP data, the details like salt tolerance or high dry matter yield is unnecessary. So better to concise portion from 96-114 in single sentence.

ANSWER:

Similarly details about imputation methods can be concise (page 7 to 12) and more details can be provided as supplementary.

ANSWER: details on FILLIN were added; for the other method, we tried to make it more coincise, when possible

Results:
--------
Need to avoid words like close to 100% better to be very precise.

ANSWER:

Maximum missing allowed and maximum percentage of missing actually present are two different things and it need to be very clear.

ANSWER: [do not know exactly what to do; leave for later on, once a new draft of the paper is complete]

Discussion
----------
Subtitles are not good - for instance "Role of crop species" and "Size of the problem".

ANSWER: some subtitles in the Discussion were changed

Previous publication demonstrating imputation in alfalfa, and other species without genome need to be discuss. [can be expanded]
Also it will be better to highlight importance of percent study while discussing previous paper comparing different imputation methods. [percent study??]

Fig 2. and 3 - Y axis need to be zoomed to make difference visible. [when possible; contrasts with reviewer 2]

ANSWER: Figures 2 and 3 were modified. One reviewer was asking to keep the y-axis constant over all comparisons, at the expense of visualizing differences between methods. The other reviewer asked to zoom on the y-axis to make such differences visible. We decide to follow the suggestion of the latter reviewer, and zommed on the y-axis; this was specified in the text and in the Figure captions, to help the reader not get confused.

I suggest to reframe the MS to make it pursuable  for readers of Molecular Breeding

ANSWER: we made an effort in that direction

Reviewer 2

Manuscript: Marker imputation efficiency for Genotyping-By- Sequencing data of crop species with and without a reference genome
Authors: Nelson Nazzicari, Filippo Biscarini, Paolo Cozzi, E. Charles Brummer, Paolo Annicchiarico

General comments
This manuscript deals with an important and emerging topic in plant (and animal) breeding as GBS is considered as a viable alternative to SNP array-based genotyping. These GBS data come with major challenges, also on the imputation of missing values. This manuscript describes the application of six algorithmic approaches to data on two plant crops that are on the edges of the wide range of genotype complexity in plants. Not surprisingly, the Beagle software performs very well (superior over all other methods) when map information is available, and much worse, when map information is absent. 
My main concerns on the paper are:

1. Why these two species? They have very different properties which makes it difficult to associate the differences in results to particular characteristics. For example, is ploidy level more critical than a reference genome?  It might have been more insightful when including an outbred diploid species with a reference genome and similar to the rice to assess imputation accuracy with and without reference genome information. [can we include other data?]

2. A missed opportunity to study the utilization of a reference genome of a related species, as is available for alfalfa

ANSWER: In alfalfa (Medicago sativa), where no reference genome is available yet, SNP markers were aligned to the genome of a closely related species, Medicago truncatula. Results have been added to the plots. [results in the text: Nelson? Io?]

3. The choice of the five imputation methods. Why these?

ANSWER:

4. Why performs Beagle still well with the unordered SNPs in the rice dataset. How severe was the re-ordering?

ANSWER: Beagle performs still relatively well in rice even when reshuffling (changing the order) the SNPs. However, imputation accuracy was 0.99 with Beagle, and 0.94-0.96 when reshuffling markers. In the minority class, it was 0.96-0.98 with ordered data, and 0.60-0.80 with reshuffled data. Reshuffling was completely random, in order to break the order of SNP along the genome. The fact that imputation was not complete mayhem, is probably due to the fact that reshuffling was performed within-chromosome, thereby preserving some LD structure. As a matter of fact, results from the entire rice dataset were worse: 0.92-0.94 total imputation accuracy and 0.43-0.64 accuracy in the minority class, with reshuffled data.

5. Too much attention for computational issues while this was not the main objective of the study (there was no effort to optimize implementation of the various methods, except for RF).

ANSWER: This section has been re-written, shortened where possible, and a statement, in the discussion, on the limitations of the information on computation time was added. [Need to add results for FILLIN --> Nelson made the plot]

6. No attention to results on individual rice chromosomes. In case there were no clear differences between chromosomes, then that should be stated more clearly.

ANSWER:

In addition, various specific comments are provided below that may be of help to further strengthen the manuscript.­

Specific comments
Line 52: It would be nice if you could provide a list of associated packages, AND explain why Beagle was selected out of these methods.

ANSWER

Line 78: There is no development of the best method in this study, nevertheless the results of this study may identify the essential properties of such a method. Was this discussed?

ANSWER:

Line 83: For alfalfa one could use the assembly of Medicago Trancatula – was this considered?

ANSWER:

Line 86: Why these four general methods?

ANSWER:

Line 109: How to interpret the 437 plants from 391 accessions as these accessions are supposed to be fully homozygous? Were there duplicates?

ANSWER: yes, they're duplicates

Line 134: Why were these values (4 and 11) were chosen? I do not see the rational of taking 4 for a heterozygote and 11 for homozygote - shouldn’t these be the reverse?

ANSWER:

Line 135-139: The reduction from 5 genotypic classes to 3 genotypic classes in alfalfa is a critical simplification step. There is no justification why this step was executed. My assumption is that this is because of the (software) methods that were used. The explanation and justification should be provided here. I am wondering whether the four general methods could be applied to the 5 genotypic classes – was this considered/studied?

ANSWER:

Line 166:  How to interpret the ‘K closest markers’ in case there is no map or assembly information (as here for alfalfa)?

ANSWER: the distance in KNN is not the physical distance between markers along the genome, but the distance between samples (plants) based on some function of the predictors (SNP genotypes, in this case). The Simple Matching Coefficient distance (Schwender, 2007) was used in this implementation of KNN.

Line 169: How was distance specified, i.e., in bp or cM?  Did you try both and if so, where there differences?

ANSWER: again, this is not the phyisical distance between markers, but the distance between observations based on a function of the SNPs (see above)

Line 173-181: I wonder whether this explanation is appropriate for this audience.

ANSWER: we made it a bit shorter, but not too much so; we think it is helpful for the reader to be reminded at least of the basic steps of a method

Line 175: Were the eigenvectors of M’M and MM’ merged to arrive at the ‘first k eigenvectors’?

ANSWER: the first k eigenvectors were determined based on the ordered corresponding eigenvalues. The matrix U* contained these k eigenvector

Line 178: Which value for k was used, was there a decision criteria used?

ANSWER: [che diciamo qui? Seguiamo Rutkoski (o che per lei)?]

Line 182: For the SVD, how critical is the initial imputation step by MNI?  Were other methods considered for this initial imputation? [I think this is just an initialization prior, and the least informative the better]

ANSWER: the initial imputation step in SVD is just aimed at providing a sort of initialization "prior", since singular value decomposition can not be performed with missing values in the matrixes. The least informative this "prior" (e.g. MNI), the better. No other methods to initialize the marker matrix for SVD were tried.

Line 185 – 198: Explain the ‘hat’ for U in equation 2. Also, why has Y no hat while there is a hat on the Y in equation 3.

ANSWER: the hat on U was a mistake and therefore removed; it is not a matrix of estimated values. However, we used the superscript U*, since this is the matrix of the first k eigenvectors of U, and therefore it is not equal to U. In equation [2], Y is the vector of observations (no hat); in equation [3] Y is the vector of predicted values (hat).

Line 217: Elaborate on the content of this likelihood file or provide a reference (to the manual of Beagle)?

ANSWER: more details were added to the text, plus a reference to the Beagle Manual

Line 229: What was the reason to not go beyond 20% artificially induced missing values?

ANSWER: we allowed maximum 70% missing genotypes per marker; in addition, we added up to 20% artificially missing genotypes. In this way, for some markers we may arrive to 90% missing data, which we considered to be the limit we didn't want to break in this study.

Line 254: It is not clear whether the ‘total number of missing data’ refers to the sum of ‘natural’ and ‘artificial’ missing data or only to the latter part. Please improve description.

ANSWER: it refers to the total number of artificially (introduced) missing genotypes. This was specified in the text.

Line 255: It may be confusing to use y in equation 4 as a discrete variable since Y is continuous in equations  [2] and [3]. It seems more intuitive to use g instead of y.

ANSWER: changed to g

Line 281/Table 1 & Table 2:  Why not merge the content of these two tables?  Transpose Table 1 to have the same columns. Then ADD the genotype frequencies for the four call-rate scenarios (10,20,40,70%).  In that way the table may reveal shifts in frequencies when imposing stronger thresholds, which may invoke further inferences on results of imputation accuracies.

ANSWER:

Line 282: Why not present these rates as percentages as that is probably easier to read and more consistent (with e.g., line 286).

ANSWER: this was done, and modified in the text

Line 284: Figure S1 does not contain information at the ‘per individual’ level

ANSWER

Line 304: How sensitive are the results to the imposed quality filters?

ANSWER:

Line 313/Figure 3: Suggest to maintain consistency in y-axis scale, i.e., the plots for Minor homozygous accuracy should range from 0 to 100% 
as well.

ANSWER: Reviewer 1 had the opposite suggestion, which we followed. See answer above.

Line 324: the flat response seems to hold for the studied range (0.01 – 0.20). Were higher values of missing rates studied?  It will be interesting to study whether there will be a critical point and whether these are different for the different genotypic classes.

ANSWER: [diplomatic answer to be added]

Line 347: Rather extensive reporting on computation time which seems not justified as the used methods were used as these were available to the authors, i.e.., there was not dedicated efforts to speed up the implemented methods.

ANSWER: This section has been re-written, shortened where possible, and a statement, in the discussion, on the limitations of the information on computation time was added. [Need to add results for FILLIN --> Nelson made the plot]

Line 379-381: Where was this negative impact shown?  In this study or other studies – please include references.

ANSWER: a reference to Rutkoski et al was added;we tried to explain better the difference between SNP array (~5% missing) and GBS data (~50% missing)

Line 401-407: This is still rather descriptive without any postulation whether any of the methods are less or more sensitive to MAF and/or balance.  The superior results of Beagle in rice are mostly driven by utilizing map/assembly information which is ignored by other methods.

ANSWER

Line 440: The reasoning on factor ii) is not clear – would the results for rice be less accurate when analysing the genome at once? Also, why would it simplify the imputation process?

ANSWER: First, the per-chromosome analysis simplifies calculations, and allow for straightforward parallelization. Additionally, the results with rice as a whole are less accurate when markers are reshuffle [add numbers/supplementary Table/Figure?]; this is because the LD structure is better preserved when reshuffling markers.

Line 443: Did you explore the population structure of the three datasets? Would presence/absence of (severe) population structure affect imputation accuracy?

Answer: i) no structure in alfalfa, presence of structure in rice (indica/tropical/temperate); ii) implicitely exploited by most imputation methods (e.g. KNN --> distance based on genetic relatedness); iii) not likely to affect much imputation accuracy; iv) this is not GWAS: the interest is not in estimating the effect of the SNP (which may be inflated if other factors are not included in the model), but in "prediction/classification" of the missing genotype to be imputed;

Line 450: It is too easy to state that there is additional complexity to make use of a reference genome of a related species. It would be of HIGH relevance to explore this as it may be applied in many species that lack a reference genome themselves.

ANSWER:

Line 535-540: Too general/vague and thus not clear why this paragraph is included in the conclusion section (or even in the manuscript)! [

ANSWER: To be checked when the new draft is ready.

Figure 1: Be consistent in using rates versus percentages, i.e., this Figure uses rates while on Line 311 you refer to percentages.

ANSWER : to be checked